---
title: "Analysis of BPC 2015  Marine Pestice Data"
output: html_notebook
---

#Libraries
```{r}
library(tidyverse)
library(magrittr)
```

#Load 2015 Sediment Data
```{r}
the.2015.data <- read.delim('BPC 2015 Sediment Samples.txt', sep = '\t')
# Note to allow  numeric input I replaced two entries (n excel) with "< 0.1" with 0.05.
the.2015.data <- select(the.2015.data, 1:15)
the.2015.data <- filter(the.2015.data, ! is.na(Location))
summary(the.2015.data)
```

#Import data on non-detect levels for those things that have NOT been detected
nd.data <- 
```{r}
library(readxl)
dl.data <- read_excel("2015 stormwater sampling DLs for Curtis ccb edits.xlsx", 
    sheet = "ND levels", skip = 2)
View(dl_data)
```

As expected, organic matter is correlated with proportion of fines in the sample
```{r}
the.2015.data %$% cor(Clay, TOC_Pct)
the.2015.data %$% cor((Clay+Silt), TOC_Pct) 
the.2015.data %$% cor((Clay)/(Clay+Silt+Sand), TOC_Pct)  # in case outliers reflect presence of carse particles

```



```{r}
plt <- ggplot(the.2015.data) + aes((Clay+Silt), TOC_Pct) + geom_point() + geom_smooth(method = 'lm')
plt

```

Bifenthrin clearly shows a pattern with TOC

```{r}
tmp <- the.2015.data %>% replace_na(list(Bifenthrin_raw = 0.045))  #VALUE NOT CONFIRMED FOR 2015 DATA
plt <- ggplot(tmp) + aes(TOC_Pct, Bifenthrin_raw) + geom_point() + geom_smooth(method = 'lm') + geom_hline(yintercept =0.045, color = 'red')
plt
rm(tmp)

```
The problem, of course, is what to do with those non-detects in estimating patterns.
Assuming the ND are zero leads to inflated variances.
```{r}
tmp <- the.2015.data %>% replace_na(list(Bifenthrin_raw = 0))
plt <- ggplot(tmp) + aes((Silt+Clay), Bifenthrin_raw) + geom_point() + geom_smooth(method = 'lm') + geom_hline(yintercept =0.045, color = 'red')
plt
rm(tmp)

```

```{r}
plt <- ggplot(the.2015.data) + aes(Bifenthrin_raw) + geom_histogram(binwidth = .1) + geom_vline(xintercept = 0.045, color = 'red')
plt
```

```{r}
plt <- ggplot(the.2015.data) + aes(Bifenthrin_raw) + geom_density() + geom_vline(xintercept = 0.045, color = 'red')
plt
```


```{r}
plt <- ggplot(the.2015.data) + aes(Bifenthrin_OC, ..density..) + geom_histogram( binwidth = 7.5) + geom_density()
plt
```





```{r}
cor(the.2015.data[4:12], use="pairwise")
cor(log(the.2015.data[4:12]), use="pairwise")
```

so pattern shows moisture is correlted with silt, and to a lesser extent, clay. WOould it be worth developing a PCR score for teh first four varibles?

```{r}
plt <- ggplot(the.2015.data, aes(Silt, Clay)) + geom_point(aes( color = Moisture_Pct), size =4) + scale_x_log10() + scale_y_log10()
plt
```

So, problem with some of this is that we use some of the physical environment values to calculate the 1*(.9/(1-.9)adjusted Bifenthrin work.  
Bifenthrin DW is adjusted for moisture content.
DW = Raw (Mass/ wet weight) x (Wet Weight/Dry Weight) = Raw   x (Mosture_pct/(1- moisture pct))

```{r}
plt <- ggplot(the.2015.data) + geom_point(aes(Moisture_Pct, Bifenthrin_raw, color = 'Raw')) +
  geom_smooth(aes(Moisture_Pct, Bifenthrin_raw, color = 'Raw'), method = 'lm') +
  geom_point(aes(Moisture_Pct, Bifenthrin_dw, color = 'Dry Weight')) +
  geom_smooth(aes(Moisture_Pct, Bifenthrin_dw, color = 'Dry Weight'), method = 'lm')+
  scale_y_log10()
plt

```
a log basis
Correction for moisture INCREASES both slope and variation. (That's more ovbious is we don't put the y scale on )


```{r}
plt <- ggplot(the.2015.data) + geom_point(aes(Moisture_Pct, Bifenthrin_raw, color = 'Raw')) +
  geom_smooth(aes(Moisture_Pct, Bifenthrin_raw, color = 'Raw'), method = 'lm') +
  geom_point(aes(Moisture_Pct, Bifenthrin_OC, color = 'Per Unit OC')) +
  geom_smooth(aes(Moisture_Pct, Bifenthrin_OC, color = 'Per Unit OC'), method = 'lm') +
  scale_y_log10()
plt

```
But correctcoion for OC reduces that again.

```{r}
cor(the.2015.data[,c(12, 4, 5, 6, 7)], use = 'pairwise')
```
the weirdness here is that the result is positively correlaated is weakly with sand

Question:  How important is the "Coarse" fraction, which is unlikely to be picked un in analytic samples.

```{r}
plt <- ggplot(the.2015.data) + aes(Coarse) + geom_histogram( bins = 10)
plt
```


```{r}
cor(the.2015.data[,c(10, 4, 5, 6, 7, 11, 9, 12,8)], use = 'pairwise')
```

Raw bifenthrin is negatively correlated with the Coarse fraction.
Bifenthrin DW is POSITIVELY correlated with moisture, which reflects the mathematical process of calculating that 

```{r}
MockWW <- seq(0,0.5,by = 0.1)  # mock values of raw concentration
#
calccDW10 <- MockWW * 100/(100-10)  #This is the functional correcton between 
calccDW30 <- MockWW * 100/(100-30)
calccDW50 <- MockWW * 100/(100-50)

test <- data.frame(MockWW, calccDW10, calccDW30, calccDW50)

plt <- ggplot(test) + aes(x = MockWW) +
  geom_line(aes(y = calccDW10, color = '10')) +
  geom_line(aes(y = calccDW30, color = '30')) + 
  geom_line(aes(y = calccDW50, color = '50')) 

plt

rm(test, MockWW, calccDW10, calccDW30, calccDW50)
```


Better wat=y to show that is the show how estimate of value in dry weight terms 

```{r}

moisture <- seq(0,0.5, by = 0.01)

wet <-  0.05

plt <- ggplot(,aes(moisture, (wet / (1-moisture)))) +
  geom_line() +
  geom_hline(yintercept = 0.05, lty = 2, color = 'red') +
  ylab('Estimated Concentration on a Dry Weight Basis')
  xlab('simulated Weght Weight Basis')
plt


```

